---
title: "HW3 - Visualization"
author: "Holly Finertie"
date: "10/14/2019"
output: github_document
---


# Problem1

```{r}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```


### Description of Instacart Data Set: 

This data set contains ``r instacart %>% nrow()` observations` and ``r instacart %>% ncol()` variables` describing order information like time and day order was placed, products ordered, aisle where products are located, and days since last order was placed per user id. For example, the individual with user id 5 ordered `r instacart %>% filter(user_id == 5) %>% summarize(max(add_to_cart_order))` items of which most were from the `r instacart %>% filter(user_id == 5) %>% summarize(max(department))` department. They placed this order `r instacart %>% filter(user_id == 5) %>% summarize(median(days_since_prior_order))` days after their last order. 

In total, there are ``r instacart %>% select(aisle) %>% n_distinct()` aisles` and the most ordered items are from the ``r instacart %>% count(aisle,name="n_aisle") %>% mutate(max_aisle = min_rank(desc(n_aisle))) %>% filter(max_aisle == 1) %>% summarize(aisle)` aisle`. 


### Plot of Items Ordered

```{r}
plot_aisles = instacart %>% 
  count(aisle, name = "n_aisle") %>% 
  filter(n_aisle > 10000) %>% 
  arrange((n_aisle)) %>% 
  ggplot(aes(x = reorder(aisle, -n_aisle), y = n_aisle)) +
  geom_bar(stat = "identity") + 
  geom_text(aes(label = n_aisle), hjust = -0.05, size = 1.5) +
  labs(
    title = "Number of Items Ordered in Aisles",
    x = "Aisle Name",
    y = "Total Items Ordered",
    caption = "Data from instacart") +
  scale_y_continuous(
    breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000), 
    labels = c("0", "25,000", "50,000", "75,000", "100,000", "125,000", "150,000"),
    limits = c(0, 160000)) +
  theme(text = element_text(size = 7))

plot_aisles + coord_flip()
```

### Most Popular Items in Each Aisle Row

```{r}
top3_products = instacart %>% 
  select(aisle, product_name) %>% 
  filter(
    aisle == "baking ingredients" | 
    aisle == "dog food care" | 
    aisle == "packaged vegetables fruits") %>% 
  group_by(aisle) %>% 
  count(product_name, name = "n_product") %>% 
  filter(min_rank(desc(n_product)) < 4) %>% 
  arrange(desc(n_product)) %>%
  mutate(
    product_name = str_to_lower(product_name), 
    "Rank" = row_number()) %>%
  rename(
    "Aisle Name" = aisle,
    "Product Name" = product_name, 
    "Count" = n_product
  ) %>% 
  select( "Rank", everything()) %>% 
  knitr::kable()

top3_products
```


### Mean Time of Day Pink Lady Apples and Coffee Ice Cream Are Ordered by Day

```{r}
apples_and_cream = instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(order_dow, product_name) %>% 
  summarize(mean_time = mean(order_hour_of_day)) %>% 
  separate(
    mean_time, 
    into = c("hour", "minute"), sep = 2) %>%
  ungroup(order_dow) %>% 
  mutate(order_dow = recode(order_dow, 
         `0` = "Monday", 
         `1` = "Tuesday", 
         `2` = "Wednesday", 
         `3` = "Thursday",
         `4` = "Friday", 
         `5` = "Saturday",
         `6` = "Sunday"), 
      hour = as.numeric(hour),
      minute = as.numeric(minute),
      ave_hour = ifelse(hour>12,(hour-12), hour),
      minute = round((minute*60), digits = 0), 
      mean_time = 
        ifelse((hour>12), paste(ave_hour,":",minute, "pm"), 
      paste(ave_hour, ":",minute, "am"))) %>% 
  select(order_dow, product_name, mean_time) %>% 
  pivot_wider(
    names_from = "product_name", 
    values_from = "mean_time"
  ) %>% 
  rename("Day of Week" = order_dow) %>% 
  knitr::kable()
apples_and_cream
```

# Problem 2

### Data Cleaning 

```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss_smart2010 = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair","Good", "Very good", "Excellent")))
```

### States with 7 or More Observed Locations (2002 and 2010)

```{r}
states_2002 = brfss_smart2010 %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  summarize(n_location = n_distinct(locationdesc)) %>% 
  filter(n_location >= 7) %>% 
  rename(
    "State" = locationabbr, 
    "Number of Locations" = n_location
  ) %>% 
  knitr::kable()

states_2002
```

* Connecticut, Florida, Massachusetts, North Carolina, New Jersey and Pennsylvania had 7 or more locations in 2002. 

```{r}
states_2010 = brfss_smart2010 %>% 
  filter(year == 2010) %>% 
  group_by(locationabbr) %>% 
  summarize(n_location = n_distinct(locationdesc)) %>% 
  filter(n_location >= 7) %>% 
  rename(
    "State" = locationabbr, 
    "Number of Locations" = n_location
  ) %>% 
  knitr::kable()

states_2010
```

* California, Colorado, Florida, Massachusetts, Maryland, North Carolina, Nebraska, New Jersey, New York, Ohio, Pennsylvania, South Carolina, Texas, and Washington had 7 or more locations in 2010. 

### Spaghetti Plot: Average Data Value Over Time by State

```{r}
brfss_excellent = brfss_smart2010  %>% 
  filter(topic == "Overall Health" & response == "Excellent") %>% 
  group_by(year, locationabbr) %>% 
  mutate(
    mean_value = mean(data_value, na.rm = TRUE)) %>% 
  select(year, locationabbr, mean_value) %>% 
  distinct() %>% 
  ggplot(aes(x = year, y = mean_value)) +
  geom_line(aes(group = locationabbr, color = locationabbr)) +
  labs(
    title = "Average Data Value for Excellent Responses",
    x = "Year",
    y = "Average Data Value",
    caption = "Data from BRFSS") +
  scale_color_hue(name = "State")

brfss_excellent
```


### 2 Panel Plot for Distribution of Data Values in NY by 2006 and 2010

```{r}
brfss_ny_state = brfss_smart2010 %>% 
  filter((year == 2010 | year == 2006) & locationabbr == "NY") %>% 
  mutate(locationdesc = substr(locationdesc, 6, 10000)) %>% 
  ggplot(aes(x = locationdesc, y = data_value, fill = response)) + 
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(~year) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(
    title = "Distribution of Data Values in NY",
    x = "NY County",
    y = "Data Value",
    caption = "Data from BRFSS") +
  scale_color_hue(name = "Response Category")
  

brfss_ny_state
```


# Problem 3

### Data Tidying

```{r}
accel_data = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "activity_minute", 
    values_to = "activity_count") %>% 
  separate(
    activity_minute, 
    into = c("activity", "activity_minute"), sep = 9) %>% 
  mutate(
    day_type = case_when(
          (day == "Saturday" | day == "Sunday") ~ "Weekend", 
          TRUE ~ "Weekday"), 
    activity_minute = as.integer(activity_minute)) %>% 
  select(-activity)
```

### Describe Data Set

This data set contains `r accel_data %>% nrow()` observations and `r accel_data %>% ncol()` variables describing five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was diagnosed with congestive heart failure (CHF). Variables include information on acitivity count `(activity_count)` for every minute `(activity_minute)` in a 24 hour period by day `(day_id)` for 5 weeks. 

### Total Activity by Day

```{r}
total_activity = accel_data %>% 
  group_by(day_id) %>% 
  summarize("Total Activity" = sum(activity_count)) %>% 
  ungroup(day_id) %>% 
  rename("Day" = day_id) %>% 
  knitr::kable()

total_activity
```

* I would argue that there are no apparent trends across the total activity for 35 days. 

### Total Activity over the course of a Day

```{r}
activity_plot = accel_data %>% 
  group_by(activity_minute, day) %>% 
  summarize(mean_activity = mean(activity_count)) %>% 
ggplot(aes(x = activity_minute, y = mean_activity)) +
  geom_line(aes(color = day), alpha = 0.4) +
  geom_smooth(se = FALSE, alpha = 0.4) +
  labs(
    title = "Activity Counts for Mintes in 24-hour Period",
    x = "Minute",
    y = "Activity Count",
    caption = "Data from the Advanced Cardiac Care 
    Center of Columbia University Medical Center"
  ) +
  scale_color_hue(name = "Day of Week") +
   scale_x_continuous(
    breaks = c(0, 180, 360, 540, 720, 900, 1080, 1260, 1440), 
    labels = c("00:00","03:00", "06:00", "09:00","12:00", "15:00", "18:00", "21:00", "24:00"),
    limits = c(0, 1440))

activity_plot
```

* The lowest activity points are around 3:00 and 24:00, while the highest activity points occur around 10:00 and 21:00 across all days. There are peaks at 10:00 on Sundays and 21:00 on Fridays close to 2,000. But on average, the activity count stays under 500. 

