---
title: "HW3 - Visualization"
author: "Holly Finertie"
date: "10/14/2019"
output: github_document
---


# Problem1

```{r}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```


### Description of Instacart Data Set

This data set "Instacart" contains ``r instacart %>% nrow()` observations` and ``r instacart %>% ncol()` variables` describing order information like time and day order was placed, products ordered, aisle where products are located, and days since last order was placed per user id. For example, the individual with user id 5 ordered `r instacart %>% filter(user_id == 5) %>% summarize(max(add_to_cart_order))` items of which most were from the `r instacart %>% filter(user_id == 5) %>% summarize(max(department))` department. They placed this order `r instacart %>% filter(user_id == 5) %>% summarize(median(days_since_prior_order))` days after their last order. 

In total, there are ``r instacart %>% select(aisle) %>% n_distinct()` aisles` and the most ordered items are from the ``r instacart %>% count(aisle,name="n_aisle") %>% mutate(max_aisle = min_rank(desc(n_aisle))) %>% filter(max_aisle == 1) %>% summarize(aisle)` aisle`. 


### Plot of Items Ordered

```{r}
plot_aisles = instacart %>% 
  count(aisle, name = "n_aisle") %>% 
  filter(n_aisle > 10000) %>% 
  arrange((n_aisle)) %>% 
  ggplot(aes(x = reorder(aisle, -n_aisle), y = n_aisle)) +
  geom_bar(stat = "identity", fill = "seagreen4") + 
  geom_text(aes(label = n_aisle), hjust = -0.05, size = 1.5) +
  labs(
    title = "Number of Items Ordered in Aisles",
    x = "Aisle Name",
    y = "Total Items Ordered",
    caption = "Data from instacart") +
  scale_y_continuous(
    breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000), 
    labels = c("0", "25,000", "50,000", "75,000", "100,000", "125,000", "150,000"),
    limits = c(0, 160000)) +
  theme(text = element_text(size = 6))

plot_aisles + coord_flip()
```

### Most Popular Items in Each Aisle Row

```{r}
top3_products = instacart %>% 
  select(aisle, product_name) %>% 
  filter(
    aisle == "baking ingredients" | 
    aisle == "dog food care" | 
    aisle == "packaged vegetables fruits") %>% 
  mutate(aisle = str_to_title(aisle)) %>% 
  group_by(aisle) %>% 
  count(product_name, name = "n_product") %>% 
  filter(min_rank(desc(n_product)) < 4) %>% 
  arrange(desc(n_product)) %>%
  mutate(
    "Rank" = row_number(),
    n_product = paste(product_name, n_product, sep = ", n = ")) %>% 
  select( "Rank", "aisle", "n_product") %>% 
  pivot_wider(
    names_from = "aisle", 
    values_from = "n_product"
  ) %>% 
  knitr::kable()

top3_products
```


### Mean Time of Day Pink Lady Apples and Coffee Ice Cream Are Ordered by Day

```{r}
apples_and_cream = instacart %>% 
  filter(
    product_name == "Pink Lady Apples" | 
    product_name == "Coffee Ice Cream") %>% 
  group_by(order_dow, product_name) %>% 
  summarize(mean_time = mean(order_hour_of_day)) %>% 
  separate(
    mean_time, 
    into = c("hour", "minute"), sep = 2) %>%
  ungroup(order_dow) %>% 
  mutate(order_dow = recode(order_dow, 
         `0` = "Sunday",
         `1` = "Monday", 
         `2` = "Tuesday", 
         `3` = "Wednesday", 
         `4` = "Thursday",
         `5` = "Friday", 
         `6` = "Saturday"), 
      hour = as.numeric(hour),
      minute = as.numeric(minute),
      minute = round((minute*60), digits = 0), 
      mean_time = paste(hour, minute, sep = ":")) %>% 
  select(order_dow, product_name, mean_time) %>% 
  pivot_wider(
    names_from = "order_dow", 
    values_from = "mean_time"
  ) %>% 
  rename("Product Name" = product_name) %>% 
  knitr::kable()

apples_and_cream
```

* Worked under the assumption that 0 = Sunday! 

# Problem 2

### Data Cleaning 

```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss_smart2010 = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair","Good", "Very good", "Excellent")))
```

### States with 7 or More Observed Locations in 2002 and 2010

```{r}
state_locations = brfss_smart2010 %>% 
  filter(year == 2002 | year == 2010) %>% 
  group_by(locationabbr, year) %>% 
  summarize(n_location = n_distinct(locationdesc)) %>% 
  filter(n_location >= 7) %>% 
  rename(
    "State" = locationabbr, 
    "Number of Locations" = n_location
  ) 
```

* In 2002, there were 6 states with 7 or more locations were: `r state_locations %>% filter(year == 2002) %>% pull(State)`. 

* In 2010, there were 14 states with 7 or more locations were: `r state_locations %>% filter(year == 2010) %>% pull(State)`. 

### Spaghetti Plot: Average Percent of Excellent Responses

```{r}
brfss_excellent = brfss_smart2010  %>% 
  filter(topic == "Overall Health" & response == "Excellent") %>% 
  group_by(year, locationabbr) %>% 
  mutate(
    mean_value = mean(data_value, na.rm = TRUE)) %>% 
  select(year, locationabbr, mean_value) %>% 
  distinct() %>% 
  ggplot(aes(x = year, y = mean_value)) +
  geom_line(aes(group = locationabbr, color = locationabbr)) +
  labs(
    title = "Average Percent of Respondents with Excellent Overall Health",
    x = "Year",
    y = "Average %",
    caption = "Data from BRFSS") +
   scale_y_continuous(
    breaks = c(10, 20, 30), 
    labels = c("10%", "20%", "30%"),
    limits = c(10, 30)) +
  viridis::scale_color_viridis(
    name = "State", 
    discrete = TRUE
  )

brfss_excellent
```


### 2 Panel Plot for Distribution of Response Types in NY by 2006 and 2010

```{r}
brfss_ny_state = brfss_smart2010 %>% 
  filter((year == 2010 | year == 2006) & locationabbr == "NY") %>% 
  mutate(
    locationdesc = substr(locationdesc, 6, 10000)) %>% 
  ggplot(aes(x = locationdesc, y = data_value, fill = response)) + 
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(~year) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(
    title = "Distribution of Responses to Overall Health in NY by County",
    x = "NY County",
    y = "% of Response Types",
    caption = "Data from BRFSS") +
   scale_y_continuous(
    breaks = c(0, 10, 20, 30, 40), 
    labels = c("0", "10%", "20%", "30%", "40%"),
    limits = c(0, 42)) +
  viridis::scale_fill_viridis(
    name = "Response Type", 
    discrete = TRUE
  )
  
brfss_ny_state
```


# Problem 3

### Data Tidying

```{r}
accel_data = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "activity_minute", 
    names_prefix = "activity_",
    values_to = "activity_count") %>% 
  mutate(
    day_type = case_when(
          (day == "Saturday" | day == "Sunday") ~ "Weekend", 
          TRUE ~ "Weekday"), 
    activity_minute = as.integer(activity_minute))
```

### Describe Data Set

This data set contains `r accel_data %>% nrow()` observations and `r accel_data %>% ncol()` variables describing five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was diagnosed with congestive heart failure (CHF). Variables include information on acitivity count `(activity_count)` for every minute `(activity_minute)` in a 24 hour period by day `(day_id)` for 5 weeks. 

### Total Activity by Day

```{r}
total_activity = accel_data %>% 
  group_by(day_id) %>% 
  summarize("Total Activity" = sum(activity_count)) %>% 
  ungroup(day_id) %>% 
  rename("Day" = day_id) %>% 
  knitr::kable()

total_activity
```

* I would argue that there are no apparent trends across the total activity for 35 days. 

### Activity by Minute by Day

```{r}
activity_plot = accel_data %>% 
  group_by(day_id, day, activity_minute) %>% 
  select(day_id, day, activity_minute, activity_count) %>% 
ggplot(aes(x = activity_minute, y = activity_count, group = day_id, color = day)) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Activity Count by Day and Minutes in 24-hour Period",
    x = "Minute",
    y = "Activity Count",
    caption = "Data from the Advanced Cardiac Care 
    Center of Columbia University Medical Center"
  ) +
  viridis::scale_color_viridis(
    name = "Day of Week", 
    discrete = TRUE
  ) +
   scale_x_continuous(
    breaks = c(0, 180, 360, 540, 720, 900, 1080, 1260, 1440), 
    labels = c("00:00","03:00", "06:00", "09:00","12:00", "15:00", "18:00", "21:00", "24:00"),
    limits = c(0, 1440)) 

activity_plot
```

* The lowest activity points for all days are from 2:00-4:00 and at 24:00, while the highest activity points occur around 10:00 on Sundays and 21:00 on Mondays, Fridays, and Saturdays. On average, the activity count stays around 500. 

